\documentclass{article}

\usepackage{listings}
\usepackage{a4}
\usepackage{courier}
\usepackage{hyperref}
\usepackage{html}


\lstdefinelanguage{ASN1} {
  morekeywords={},
  sensitive=false,
  morecomment=[s]{(--}{--)}
  }

\lstdefinelanguage{shell} {
  sensitive=true
  }

\setlength{\parskip}{\medskipamount}
\setlength{\parindent}{0pt}

\title{Haskell ASN.1 PER Library 1.0.0}
\author{Dominic Steinitz}

\begin{document}

\maketitle

The 
\htmladdnormallinkfoot
{Haskell Cryptographic Library 1.0.0}
{http://www.haskell.org/asn1}
is an implementation of the ASN.1 Packed Encoding Rules (PER).


The current release does not contain a parser for ASN.1 so you have to hand encode
ASN.1 definitions into the Haskell ASN.1 Abstract Syntax Tree (AST). Examples of these are given
for each ASN.1 construction supported.

Haddock documentation for the library is available
\htmladdnormallinkfoot
{here}
{http://www.haskell.org/asn1/doc/html}
; alternatively, you can generate it yourself (need to give instructions on how to do this).

\section{ASN.1 Support}

\subsection{BOOLEAN}

\lstset{language=Haskell}
\begin{lstlisting}[frame=single]
BOOLEAN :: ASNType Bool

falseEncoding = toPer BOOLEAN False
\end{lstlisting}

\subsection{INTEGER}

\subsection{ENUMERATED}

\subsection{REAL}

We don't currently support this. Do we want to?

\subsection{BIT STRING}

\begin{lstlisting}[frame=single]
BITSTRING :: NamedBits -> ASNType BitString

bitString = 
   toPer (BITSTRING []) 
         (BitString [1,1,0,0,0,1,0,0,0,0])
\end{lstlisting}

We don't seem to have any named bits tests or examples.

\section{Installation Instructions}

Get the sources:

\lstset{language=shell,basicstyle=\ttfamily\small}
\begin{lstlisting}[frame=single]
darcs get --tag "4.0.3" http://darcs.haskell.org/crypto
\end{lstlisting}

Build and install ready for testing:

\begin{lstlisting}[frame=single]
ghc -o Setup Setup.hs -package Cabal
./Setup configure --prefix=/my/chosen/dir
./Setup build
./Setup install --user
\end{lstlisting}

Run the tests.

\begin{lstlisting}[frame=single]
cd /my/chosen/dir/bin
./RSATest
./SymmetricTest
./QuickTest
\end{lstlisting}

You can now run the examples to confirm further that everything
is working satisfactorily.
When you are happy, build and install them in
their final destination:

\begin{lstlisting}[frame=single]
./Setup unregister --user
./Setup clean
./Setup configure
./Setup build
./Setup install
\end{lstlisting}

\section{To Do}

In no particular order:

\begin{itemize}
\item
Incorporate other symmetric key algorithms already coded in Haskell.
\item
Performance analysis as Blowfish ought to run more quickly than DES.
\item
Other modes / padding schemes.
\item
Extend typechecking to ensure that only the appropriate key sizes are used
for a given algorithm.
\item
Improve performance, for example, for SHA1. This 
\htmladdnormallinkfoot
{code}
{http://www.abridgegame.org/repos/darcs-unstable}
runs an order of magnitude faster but, with respect to the authors, doesn't
feel that functional.
\item
Get rid of the GPL code.
\end{itemize}

\section{Contact}

All questions, comments, bug reports, flames, requests for 
updates / changes and suggestions should be directed to Dominic Steinitz and
logged
\htmladdnormallinkfoot
{here}
{http://hackage.haskell.org/trac/crypto}
.
\section{Licensing}

The modules in the library come from different authors and have been 
released under different licences. 

\subsection{Contributors}

\subsubsection{Codec.Binary}
\begin{tabular}{|p{6cm}|p{3cm}|p{3cm}|p{1cm}|}
\hline\hline
Codec.Binary.Base64 & Warrick Gray & 
Copyright \copyright\ 2005, All rights reserved & BSD \\
\hline\hline
\end{tabular}

\subsubsection{Codec.Text}
\begin{tabular}{|p{6cm}|p{3cm}|p{3cm}|p{1cm}|}
\hline\hline
Codec.Text.Raw & Dominic Steinitz & 
Copyright \copyright\ 2006, All rights reserved & BSD \\
\hline\hline
\end{tabular}

\subsubsection{Codec.Encryption}
\begin{tabular}{|p{6cm}|p{3cm}|p{3cm}|p{1cm}|}
\hline\hline
Codec.Encryption.AES & Lukasz Anforowicz &
Copyright \copyright\ 2005, All rights reserved & BSD \\
\hline
Codec.Encryption.AESAux & Dominic Steinitz &
Copyright \copyright\ 2005, All rights reserved & BSD \\
\hline
Codec.Encryption.Blowfish & Doug Hoyte &
Copyright \copyright\ 2005, All rights reserved & BSD \\
\hline
Codec.Encryption.BlowfishAux & Dominic Steinitz &
Copyright \copyright\ 2005, All rights reserved & BSD \\
\hline
Codec.Encryption.DES & Ian Lynagh &
Copyright \copyright\ 2005, All rights reserved & BSD \\
\hline
Codec.Encryption.DESAux & Dominic Steinitz &
Copyright \copyright\ 2005, All rights reserved & BSD \\
\hline
Codec.Encryption.Modes & Dominic Steinitz &
Copyright \copyright\ 2005, All rights reserved & BSD \\
\hline
Codec.Encryption.Padding & Dominic Steinitz &
Copyright \copyright\ 2005, All rights reserved & BSD \\
\hline
Codec.Encryption.RSA & David Sankel &
Copyright \copyright\ 2005, All rights reserved & GPL \\
\hline
Codec.Encryption.RSA.EMEOAEP & David Sankel &
Copyright \copyright\ 2005, All rights reserved & GPL \\
\hline
Codec.Encryption.RSA.MGF & Dominic Steinitz &
Copyright \copyright\ 2005, All rights reserved & BSD \\
\hline
Codec.Encryption.RSA.NumberTheory & David Sankel &
Copyright \copyright\ 2005, All rights reserved & GPL \\
\hline\hline
\end{tabular}

\subsubsection{Codec}
\begin{tabular}{|p{6cm}|p{3cm}|p{3cm}|p{1cm}|}
\hline\hline
Codec.Utils & Dominic Steinitz &
Copyright \copyright\ 2005, All rights reserved & BSD \\
\hline\hline
\end{tabular}

\subsubsection{Data.Digest}
\begin{tabular}{|p{6cm}|p{3cm}|p{3cm}|p{1cm}|}
\hline\hline
Data.Digest.MD5 & Dominic Steinitz &
Copyright \copyright\ 2005, All rights reserved & BSD \\
\hline
Data.Digest.MD5Aux & Ian Lynagh &
Copyright \copyright\ 2005, All rights reserved & BSD \\
\hline
Data.Digest.SHA1 & Dominic Steinitz &
Copyright \copyright\ 2005, All rights reserved & BSD \\
\hline\hline
\end{tabular}

\subsubsection{Data}
\begin{tabular}{|p{6cm}|p{3cm}|p{3cm}|p{1cm}|}
\hline\hline
Data.LargeWord 
& Dominic Steinitz &
Copyright \copyright\ 2005, All rights reserved & BSD \\
\hline\hline
\end{tabular}

\subsubsection{Tests and Examples}
\begin{tabular}{|p{6cm}|p{3cm}|p{3cm}|p{1cm}|}
\hline\hline
RSATest
& Dominic Steinitz &
Copyright \copyright\ 2005, All rights reserved & BSD \\
\hline
QuickTest
& Dominic Steinitz &
Copyright \copyright\ 2005, All rights reserved & BSD \\
\hline
SymmetricTest
& Dominic Steinitz &
Copyright \copyright\ 2005, All rights reserved & BSD \\
\hline\hline
\end{tabular}

\subsection{The BSD License}

This license is based on
\htmladdnormallinkfoot
{The BSD License}
{http://www.opensource.org/licenses/bsd-license.php}.

Redistribution and use in source and binary forms, with or without 
modification, are permitted provided that the following conditions are met:

\begin{itemize}
\item
Redistributions of source code must retain the above copyright notice, 
this list of conditions and the following disclaimer.
\item
Redistributions in binary form must reproduce the above copyright notice, 
this list of conditions and the following disclaimer in the documentation 
and/or other materials provided with the distribution.
\item
The names of its contributors may not be used to endorse or promote 
products derived from this software without specific prior written permission.
\end{itemize}

\begin{sc}
This software is provided by the copyright holders and contributors ``AS IS'' 
and any express or implied warranties, including, but not limited to, 
the implied warranties of merchantability and fitness for a particular 
purpose are disclaimed. In no event shall the copyright onwers or
contributors be liable for any direct, indirect, incidental, special,
exemplary, or consequential damages (including, but not limited to,
procurement of substitute goods or services; loss of use, data, or profits;
or business interruption) however caused and on any theory of liability,
whether in contract, strict liability, or tort (including negligence or
otherwise) arising in any way out of the use of this software,
even if advised of the possibility of such damage.
\end{sc}

\subsection{The GNU General Public License (GPL)}

This program is free software; you can redistribute it and/or modify it under the terms of the GNU General Public License as published by the Free Software Foundation; either version 2 of the License, or (at your option) any later version.

This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License for more details.

You can find a copy of
the GNU General Public License 
\htmladdnormallinkfoot
{here}
{http://www.opensource.org/licenses/gpl-license.php}
; 
if not, write to the Free Software Foundation, Inc., 59 Temple Place, 
Suite 330, Boston, MA 02111-1307 USA

\section{Disclaimer}
Cryptography is a notoriously easy area in which to make mistakes, 
not necessarily with the algorithms but with how they are implemented 
(for example not protecting keys, using weak keys and so on). 
For a readable account of some of the pitfalls, see 
\htmladdnormallinkfoot
{Ross Anderson}
{http://www.cl.cam.ac.uk/users/rja14/}
's book.

\begin{sc}
This software is provided by the copyright holders and contributors ``AS IS'' 
and any express or implied warranties, including, but not limited to, 
the implied warranties of merchantability and fitness for a particular 
purpose are disclaimed. In no event shall the copyright onwers or
contributors be liable for any direct, indirect, incidental, special,
exemplary, or consequential damages (including, but not limited to,
procurement of substitute goods or services; loss of use, data, or profits;
or business interruption) however caused and on any theory of liability,
whether in contract, strict liability, or tort (including negligence or
otherwise) arising in any way out of the use of this software,
even if advised of the possibility of such damage.
\end{sc}

\section{Acknowledgements}

\begin{itemize}
\item
Doug Hoyte (HardCore SoftWare)
\item
Anatoly Zaretsky
\item
\htmladdnormallinkfoot
   {Ian Lynagh}
   {http://web.comlab.ox.ac.uk/oucl/work/ian.lynagh}
\item
\htmladdnormallinkfoot
   {David Sankel}
   {http://www.electronconsulting.com/whois.html}
\item
\htmladdnormallinkfoot 
{Ross Paterson}
{http://www.soi.city.ac.uk/~ross}
\item
Lukasz Anforowicz
\item
\htmladdnormallinkfoot 
{Warrick Gray}
{http://homepages.paradise.net.nz/warrickg/haskell/http/}
\item
\htmladdnormallinkfoot 
{Russell O'Connor}
{http://r6.ca}
\item
Spencer Janssen
\end{itemize}

This document was last updated on 1st May 2007.
\copyright\ 2006--2007 Dominic Steinitz. 

\end{document}
